---
title: "Vincent + Raymond Second Attempt"
output: html_notebook
---

# Load required packages

SIAMCAT is used for analysis

```{r}
library(curatedMetagenomicData)
library(dplyr)
library(mia)
library(SIAMCAT)
library(stringr)
library(ggplot2)
```

# Filter cMD metadata table for Vincent and Raymond
```{r}
meta <- curatedMetagenomicData::sampleMetadata
meta$study_name %>% table()

submeta <- meta %>% filter(study_name=="VincentC_2016" | study_name=="RaymondF_2016")
submeta$study_name %>% table()
```

# Download data from cMD for matching metadata
Table should match the metadata table
```{r}
rayvince <- returnSamples(submeta, "relative_abundance")
colData(rayvince)$study_name %>% table()
```

# Proportions in cMD are presented as percentages. Need to be decimals. Divide by 100.
```{r}
assay(rayvince) <- assay(rayvince)/100
```

#Subset to just Raymond post-exposure cases

There should be 24 data points (18 cases, 6 controls)
```{r}
ray <- subset(rayvince, select =colData(rayvince)$study_name=="RaymondF_2016")
#ray <- subset(rayvince, select =colData(rayvince)$study_name=="RaymondF_2016" &
              #  colData(rayvince)$days_from_first_collection == 7)

colData(ray)$study_name %>% table()
```
```{r}
phyloray <- makePhyloseqFromTreeSummarizedExperiment(ray, abund_values = "relative_abundance")
```
# Extract sample and rel abundance tables

Not sure that this is necessary. I think SIAMCAT can just read the phyloseq object.

```{r}
sample.phyloray <- sample_data(phyloray)
feat.phyloray <- otu_table(phyloray)
```

# Remove duplicate variables

Remove two variables that are identical with condition of interest as they were causing issues down the line

```{r}
sample.phyloray <- sample.phyloray[, -which(names(sample.phyloray) %in% c("antibiotics_family", "antibiotics_current_use"))]

```

# Set variable of interest

```{r}
label.ray <- create.label(meta=sample.phyloray, label="study_condition", case="cephalosporins")
```

# Create SIAMCAT object

```{r}
sc.obj <- siamcat(feat=feat.phyloray, label=label.ray, meta=sample.phyloray)
```

# Filter by abundance

Cutoff is set at 0.001. Might want to experiment with this.

```{r}
sc.filt <- filter.features(sc.obj, filter.method = "abundance", cutoff = 0.001)
```

# Check associations

```{r}

sc.assc <- check.associations(
  sc.filt,
  sort.by = 'fc',
  alpha = 0.05,
  fn.plot = 'assocation_plots.pdf',
  mult.corr = "fdr",
  feature.type = "filtered",
  detect.lim = 10 ^-6,
  plot.type = "quantile.box",
  panels = c("fc", "prevalence", "auroc"))
```

# Check Confounders

This checks confounders--does not work if any variable 100% matches condition of interest which is why we removed those two other variables earlier.

```{r}

sc.conf <- check.confounders(
  sc.filt,
  fn.plot = 'confounder_plots.pdf',
  meta.in = NULL,
  feature.type = 'filtered', verbose = 1
)

```

# Normalization

I'm guessing due to ranked unit normalization this is using Rankits

```{r}
sc.norm <- normalize.features(
  sc.filt,
  norm.method = "rank.unit"
  )
```

# Prepare to run machine learning task

Split the data, not sure the difference between folds and resamples and how to select the right number--right now I just picked numbers that didn't give an error message

```{r}

sc.split <-  create.data.split(
  sc.norm,
  num.folds = 6,
  num.resample = 2,
  inseparable = "subject_id"
)

```

# Random Forest

Lasso also seems to work--others gave weird results (weird ROCs)

```{r}

sc.mod <- train.model(
  sc.split,
  method = "randomForest"
)

sc.pred <- make.predictions(sc.mod)
pred_matrix <- pred_matrix(sc.pred)

```

# View ROC and PR curve

```{r}
sc.eval <- evaluate.predictions(sc.pred)
model.evaluation.plot(sc.eval)
```
# Prediction Spaghetti Plots
```{r}
library(stringr)
spag <- data.frame(pred_matrix, days=rownames(pred_matrix) %>% str_extract(pattern="[EC]\\w*") %>% str_sub(start=2) %>% as.integer(), subj = rownames(pred_matrix) %>% str_extract("[P]\\d{1,2}"))
colnames(spag) <- c("Predicted Probability", "CV_rep2", "Days", "subj")

ggplot(data=spag) + geom_line(aes(x=Days, y=`Predicted Probability`, group=subj, color=subj)) + geom_boxplot(aes(x=Days,y=`Predicted Probability`, group=Days, alpha = 1)) + theme_minimal() +  theme(legend.position="none")


```



# Export final model interpretation plot as PDF

```{r}
model.interpretation.plot(
sc.eval,
fn.plot = 'interpretation.pdf',
consens.thres = 0.01,
limits = c(-3, 3),
heatmap.type = 'zscore'
)
```

```{r}
vince <- subset(rayvince, select=colData(rayvince)$study_name=="VincentC_2016")
colData(vince)$abx <- "control"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("cephalosporins")] <- "cephalosporins"
colData(vince)$abx[!is.na(colData(vince)$antibiotics_family)] <- "cephalosporins"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("fluoroquinolones")] <- "cephalosporins"
colData(vince)$abx %>% table()
```

```{r}
phylovince <- makePhyloseqFromTreeSummarizedExperiment(vince, abund_values = "relative_abundance")
```

```{r}
sample.phylovince <- sample_data(phylovince)
feat.phylovince <- otu_table(phylovince)
```

```{r}
sample.phylovince <- sample.phylovince[, -which(names(sample.phylovince) %in% c("antibiotics_family", "antibiotics_current_use"))]
```

```{r}
label.vince <- create.label(meta=sample.phylovince, label="abx", case="cephalosporins")
```

```{r}
sc.obj.test <- siamcat(feat=feat.phylovince, label=label.vince, meta=sample.phylovince)
```

```{r}
sc.obj.filt <- filter.features(sc.obj.test, filter.method = "abundance", cutoff = 0)
sc.obj.norm <- normalize.features(sc.obj.filt, norm.param=norm_params(sc.eval))

sc.obj.test <- make.predictions(sc.eval, sc.obj.norm)
sc.obj.test <- evaluate.predictions(sc.obj.test)
model.evaluation.plot(sc.obj.test)

```

```{r}
sc.vince.split <-  create.data.split(
  sc.obj.norm,
  num.folds = 6,
  num.resample = 2,
  inseparable = "subject_id"
)

sc.vince.mod <- train.model(
  sc.vince.split,
  method = "randomForest"
)

sc.vince.pred <- make.predictions(sc.vince.mod)
pred_matrix_vince <- pred_matrix(sc.vince.pred)

sc.vince.eval <- evaluate.predictions(sc.vince.pred)
model.evaluation.plot(sc.vince.eval)

```

# Take the Vincent dataset and filter down to either the last assessment if no abx or the abx assessment

# Figure out participant IDs that were cases and controls
```{r}

library(dplyr)
dsVince <- colData(vince) %>% as.data.frame()
dsVince$abx <- 0
dsVince$abx[!is.na(dsVince$antibiotics_family)] <- 1

cases <- dsVince %>% group_by(subject_id) %>% summarize(abx=sum(abx))

```

# Genus-Level Analysis

```{r}
rayvincegenus <- agglomerateByRank(rayvince, rank="genus")
rayg <- subset(rayvincegenus, select =colData(rayvince)$study_name=="RaymondF_2016")
phyloray <- makePhyloseqFromTreeSummarizedExperiment(rayg, abund_values = "relative_abundance")
sample.phyloray <- sample_data(phyloray)
feat.phyloray <- otu_table(phyloray)
sample.phyloray <- sample.phyloray[, -which(names(sample.phyloray) %in% c("antibiotics_family", "antibiotics_current_use"))]
label.ray <- create.label(meta=sample.phyloray, label="study_condition", case="cephalosporins")
sc.obj <- siamcat(feat=feat.phyloray, label=label.ray, meta=sample.phyloray)
sc.filt <- filter.features(sc.obj, filter.method = "abundance", cutoff = 0.001)
sc.norm <- normalize.features(
  sc.filt,
  norm.method = "rank.unit"
  )
sc.norm <- normalize.features(
  sc.filt,
  norm.method = "rank.unit"
  )
sc.split <-  create.data.split(
  sc.norm,
  num.folds = 6,
  num.resample = 2,
  inseparable = "subject_id"
)
sc.mod <- train.model(
  sc.split,
  method = "randomForest"
)

sc.pred <- make.predictions(sc.mod)
pred_matrix <- pred_matrix(sc.pred)
sc.eval <- evaluate.predictions(sc.pred)
model.evaluation.plot(sc.eval)
```

# Genus-level model
```{r}
vinceg <- subset(rayvincegenus, select=colData(rayvince)$study_name=="VincentC_2016")
colData(vinceg)$abx <- "control"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("cephalosporins")] <- "cephalosporins"
colData(vinceg)$abx[!is.na(colData(vinceg)$antibiotics_family)] <- "cephalosporins"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("fluoroquinolones")] <- "cephalosporins"
phylovince <- makePhyloseqFromTreeSummarizedExperiment(vinceg, abund_values = "relative_abundance")
sample.phylovince <- sample_data(phylovince)
feat.phylovince <- otu_table(phylovince)
sample.phylovince <- sample.phylovince[, -which(names(sample.phylovince) %in% c("antibiotics_family", "antibiotics_current_use"))]
label.vince <- create.label(meta=sample.phylovince, label="abx", case="cephalosporins")
sc.obj.test <- siamcat(feat=feat.phylovince, label=label.vince, meta=sample.phylovince)
sc.obj.filt <- filter.features(sc.obj.test, filter.method = "abundance", cutoff = 0)
sc.obj.norm <- normalize.features(sc.obj.filt, norm.param=norm_params(sc.eval))

sc.obj.test <- make.predictions(sc.eval, sc.obj.norm)
sc.obj.test <- evaluate.predictions(sc.obj.test)
model.evaluation.plot(sc.obj.test)
```