---
title: "Train on Vincent, Test on Raymond"
output: html_notebook
---

# Load required packages

SIAMCAT is used for analysis

```{r}
library(curatedMetagenomicData)
library(dplyr)
library(mia)
library(SIAMCAT)
library(stringr)
```

# Filter cMD metadata table for Vincent and Raymond
```{r}
meta <- curatedMetagenomicData::sampleMetadata
meta$study_name %>% table()

submeta <- meta %>% filter(study_name=="VincentC_2016" | study_name=="RaymondF_2016")
submeta$study_name %>% table()
```

# Download data from cMD for matching metadata
Table should match the metadata table
```{r}
rayvince <- returnSamples(submeta, "relative_abundance")
colData(rayvince)$study_name %>% table()
```

# Proportions in cMD are presented as percentages. Need to be decimals. Divide by 100.
```{r}
assay(rayvince) <- assay(rayvince)/100
```


```{r}
vince <- subset(rayvince, select=colData(rayvince)$study_name=="VincentC_2016")
colData(vince)$abx <- "control"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("cephalosporins")] <- "cephalosporins"
colData(vince)$abx[!is.na(colData(vince)$antibiotics_family)] <- "cephalosporins"
#colData(vince)$abx[colData(vince)$antibiotics_family %>% str_detect("fluoroquinolones")] <- "cephalosporins"
colData(vince)$abx %>% table()
```

```{r}
phylovince <- makePhyloseqFromTreeSummarizedExperiment(vince, abund_values = "relative_abundance")
```

```{r}
sample.phylovince <- sample_data(phylovince)
feat.phylovince <- otu_table(phylovince)
```

```{r}
sample.phylovince <- sample.phylovince[, -which(names(sample.phylovince) %in% c("antibiotics_family", "antibiotics_current_use"))]
```

```{r}
label.vince <- create.label(meta=sample.phylovince, label="abx", case="cephalosporins")
```

# Create SIAMCAT object

```{r}
sc.obj <- siamcat(feat=feat.phylovince, label=label.vince, meta=sample.phylovince)
```

# Filter by abundance

Cutoff is set at 0.001. Might want to experiment with this.

```{r}
sc.filt <- filter.features(sc.obj, filter.method = "abundance", cutoff = 0.001)
```

# Normalization

I'm guessing due to ranked unit normalization this is using Rankits

```{r}
sc.norm <- normalize.features(
  sc.filt,
  norm.method = "rank.unit"
  )
```

# Prepare to run machine learning task

Split the data, not sure the difference between folds and resamples and how to select the right number--right now I just picked numbers that didn't give an error message

```{r}

sc.split <-  create.data.split(
  sc.norm,
  num.folds = 6,
  num.resample = 2,
  inseparable = "subject_id"
)

```

# Random Forest

Lasso also seems to work--others gave weird results (weird ROCs)

```{r}

sc.mod <- train.model(
  sc.split,
  method = "randomForest"
)

sc.pred <- make.predictions(sc.mod)
pred_matrix <- pred_matrix(sc.pred)

```

# Subset to just Raymond  cases

There should be 24 data points (18 cases, 6 controls)
```{r}
ray <- subset(rayvince, select =colData(rayvince)$study_name=="RaymondF_2016")
#ray <- subset(rayvince, select =colData(rayvince)$study_name=="RaymondF_2016" &
              #  colData(rayvince)$days_from_first_collection == 7)

colData(ray)$study_name %>% table()
```
```{r}
phyloray <- makePhyloseqFromTreeSummarizedExperiment(ray, abund_values = "relative_abundance")
```
# Extract sample and rel abundance tables

Not sure that this is necessary. I think SIAMCAT can just read the phyloseq object.

```{r}
sample.phyloray <- sample_data(phyloray)
feat.phyloray <- otu_table(phyloray)
```

# Remove duplicate variables

Remove two variables that are identical with condition of interest as they were causing issues down the line

```{r}
sample.phyloray <- sample.phyloray[, -which(names(sample.phyloray) %in% c("antibiotics_family", "antibiotics_current_use"))]

```

# Set variable of interest

```{r}
label.ray <- create.label(meta=sample.phyloray, label="study_condition", case="cephalosporins")
```





# View ROC and PR curve

```{r}
sc.eval <- evaluate.predictions(sc.pred)
model.evaluation.plot(sc.eval)
```


```{r}
sc.obj.test <- siamcat(feat=feat.phylovince, label=label.vince, meta=sample.phylovince)
```

```{r}
sc.obj.filt <- filter.features(sc.obj.test, filter.method = "abundance", cutoff = 0)
sc.obj.norm <- normalize.features(sc.obj.filt, norm.param=norm_params(sc.eval))

sc.obj.test <- make.predictions(sc.eval, sc.obj.norm)
sc.obj.test <- evaluate.predictions(sc.obj.test)
model.evaluation.plot(sc.obj.test)

```
